Longest Common Extensions via Fingerprinting

### 2012-03-02 article-presentation/presentation.tex
clearer example
updates based on feedback from presentation rehersal at DTU

\subsection{The LCE Problem}
\begin{frame}{The LCE Problem}
    \begin{description}
        \item[LCE value] $\LCE_s(i,j)$ is the length of the longest common prefix of the two suffixes of a string $s$ starting at index $i$ and $j$
        \item[LCE problem] Efficiently query multiple LCE values on a static string $s$ and varying pairs $(i,j)$
    \end{description}
    Example:
    \begin{description}
        \item[input:] $s=abbababba$, $(i,j)=(4,6)$
        \item[suffix $i$] = $ababba$
        \item[suffix $j$] = $abba$
        \item[longest common prefix $(i,j)$] = $ab$
        \item[$\LCE_s(i,j)$] = $2$
    \end{description}
\end{frame}

@@

\subsection{Practical results}
\begin{frame}{Existing Algorithms: Practical Results}
    Query times of
    \textcolor{red}{\proc{DirectComp}} and
    \textcolor{green}{\proc{LcpRmq}}
    by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-rmq-directcomp-slide-rand10.plt}
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-rmq-directcomp-slide-alla.plt}
        \end{center}
    }
    \end{multicols}

@@

\begin{comment}
\subsection{Preprocessing}
\begin{frame}{The \fprintk\ Algorithm: Preprocessing}
    \begin{itemize}
        \item For each level $\ell$
        \begin{itemize}
            \item For each $t_\ell$-length substring in lexicographically sorted order
            \begin{itemize}
                \item If the current substring $s[\SA[i]\twodots\SA[i]+t_\ell-1]$ is equal to the previous substring, give it the same fingerprint as the previous substring, otherwise give it a new unused fingerprint. The two substrings are equal when $\LCE[i] \geq t_\ell$.
            \end{itemize}
        \end{itemize}
    \end{itemize}

    \begin{multicols}{2}{
        \begin{center}
        \hfill\\
            $s=$ abbababba\\\hfill\\
            For $t_\ell = 3$:\\
            $H_\ell =$ [3, 8, 6, 2, 6, 3, 8, 5, 1]
        \end{center}
        \newpage
        \begin{center}
            \includegraphics[width=0.2\textwidth,page=1]{../doc/fingerprint-preproc.pdf}
        \end{center}
    }
    \end{multicols}

@@

    \begin{tabular}{r l l l}
        & $k=1$ & $k=2$ & $k=\logceil$ \\
        %Preprocessing & $O(\sortt)$ & $O(\sortt)$ & $O(n\log n)$ \\
        Space & $O(n)$ & $O(n)$ & $O(n\log n)$ \\
        Query & $O(n)$ & $O(\sqrt n)$ & $O(\log n)$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        %Query I/O & $O\Big(\frac{n}{B}\Big)$ & $O\Big(\frac{\sqrt n}{B}\Big)$ & $O(\log n)$ \\
    \end{tabular}
\end{frame}

@@

\section{Summary}
\begin{frame}{Summary}
    \begin{tabular}{r l l l}
        & \proc{Direct-} & \proc{LcpRmq} / & \\
        & \proc{Comp} & \proc{SuffixNca} & \fprintk \\
        %Preprocessing & $O(1)$ & $O(\sortt)$ & $O(k\cdot n+\sortt)$ \\
        Space & $O(1)$ & $O(n)$ & $O(k\cdot n)$\\
        Query & $O(n)$ & $O(1)$ & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        Query I/O & $O\big(\frac{n}{B}\big)$ & $O(1)$ & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
    \end{tabular}
    \vspace{1em}
    \begin{itemize}
        \item In practice, the \fprintk\ algorithm is...
        \begin{itemize}
            \item ...almost as good as \proc{DirectComp} and significantly better than \proc{LcpRmq} in average case
            \item ...significantly better than \proc{DirectComp} but worse than \proc{LcpRmq} in worst case
        \end{itemize}
        \item Cache optimization of \fprintk\ improves query times at $k=2$ and worsens query times at $k\geq 3$
    \end{itemize}
\end {frame}

>>>>

\subsection{The LCE Problem}
\begin{frame}{The LCE Problem}
    \begin{description}
        \item[LCE value] $\LCE_s(i,j)$ is the length of the longest common prefix of the two suffixes of a string $s$ starting at index $i$ and $j$
        \item[LCE problem] Efficiently query multiple LCE values on a static string $s$ and varying pairs $(i,j)$
    \end{description}
    Example:
    \begin{description}
        \item[input:] $s=abbababba$, $(i,j)=(4,6)$
        \item[suffix $i$ of $s$] = $ababba$
        \item[suffix $j$ of $s$] = $abba$
        \item[longest common prefix] = $ab$
        \item[$\LCE_s(i,j)$] = $2$
    \end{description}
\end{frame}

@@

\subsection{Practical results}
\begin{frame}{Existing Algorithms: Practical Results}
    Query times of
    \textcolor{red}{\proc{DirectComp}} and
    \textcolor{green}{\proc{LcpRmq}}
    by string length
    \begin{multicols}{2}{
        \begin{center}
            Average case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-rmq-directcomp-slide-rand10.plt}\\
            $s=$ random characters\\ $\sigma=10$
            \newpage
            Worst case
            \includegraphics[width=0.49\textwidth,type=pdf,ext=.pdf,read=.pdf]{../src/results/length-rmq-directcomp-slide-alla.plt}\\
            $s=$ aaaaa...
        \end{center}
    }
    \end{multicols}

@@

%\begin{comment}
\subsection{Preprocessing}
\begin{frame}{The \fprintk\ Algorithm: Preprocessing}
    \begin{itemize}
        \item For each level $\ell$
        \begin{itemize}
            \item For each $t_\ell$-length substring in lexicographically sorted order
            \begin{itemize}
                \item If the current substring $s[\SA[i]\twodots\SA[i]+t_\ell-1]$ is equal to the previous substring, give it the same fingerprint as the previous substring, otherwise give it a new unused fingerprint. The two substrings are equal when $\LCP[i] \geq t_\ell$.
            \end{itemize}
        \end{itemize}
    \end{itemize}

    \begin{multicols}{2}{
        \begin{center}
            \includegraphics[width=0.2\textwidth,page=1]{../doc/fingerprint-preproc.pdf}
        \end{center}
        \newpage
        %\begin{center}
        \hfill\\
            $s=$ abbababba\\\hfill\\
            For $t_\ell = 3$:\\
            $H_\ell =$ [3, 8, 6, 2, 6, 3, 8, 5, 1]
        %\end{center}
    }
    \end{multicols}

@@

    \begin{tabular}{r l l l}
        & $k=1$ & $k=2$ & $k=\logceil$ \\
        %Preprocessing & $O(\sortt)$ & $O(\sortt)$ & $O(n\log n)$ \\
        Space & $O(n)$ & $O(n)$ & $O(n\log n)$ \\
        Query & $O(n)$ & $O(\sqrt n)$ & $O(\log n)$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        %Query I/O & $O\Big(\frac{n}{B}\Big)$ & $O\Big(\frac{\sqrt n}{B}\Big)$ & $O(\log n)$ \\
    \end{tabular}

    Space for \fprintk\ is the same as for \proc{LcpRmq} when $k=6$.
\end{frame}

@@

\section{Summary}
\begin{frame}{Summary}
    \begin{tabular}{r l l l}
        & \proc{Direct-} & \proc{LcpRmq} / & \\
        & \proc{Comp} & \proc{SuffixNca} & \fprintk \\
        %Preprocessing & $O(1)$ & $O(\sortt)$ & $O(k\cdot n+\sortt)$ \\
        Space & $O(1)$ & $O(n)$ & $O(k\cdot n)$\\
        Query & $O(n)$ & $O(1)$ & $O(k\cdot n^{1/k})$ \\
        Average query & $O(1)$ & $O(1)$ & $O(1)$ \\
        & fast & slow & fast \\
        Query I/O & $O\big(\frac{n}{B}\big)$ & $O(1)$ & $O\Big(k\cdot\Big(\frac{n^{1/k}}{B}+1\Big)\Big)$ \\
        Code complexity & very simple & complex & simple \\
    \end{tabular}
    \vspace{1em}
    \begin{itemize}
        %\item In practice, the \fprintk\ algorithm is...
        %\begin{itemize}
        %    \item ...almost as good as \proc{DirectComp} and significantly better than \proc{LcpRmq} in average case
        %    \item ...significantly better than \proc{DirectComp} but worse than \proc{LcpRmq} in worst case
        %\end{itemize}
        \item Cache optimization of \fprintk\ improves query times at $k=2$ and worsens query times at $k\geq 3$
        \item Space for \fprintk\ is the same as for \proc{LcpRmq} when $k=6$.
    \end{itemize}
\end {frame}
